02 - PubliliaFitness - Defining strategies and calculating frequencies

### 2023-09-01 scripts/Publilia-02-strats.Rmd
fixed figures 2, 3, and s4

base_text <- expand.grid(ants = c("Ants removed", "Control"),
            `visited in 4 days` = c("Not visited\nwithin first 4 days", 
                                    "Visited\nwithin first 4 days")) %>%
  tibble()

hatch_text <- base_text %>%
  mutate(label = ifelse(ants=="Ants removed" & `visited in 4 days`=="Not visited\nwithin first 4 days", 
         "Hatch date", 
         ""))

(guarding_durations_plot_2023 <- strats_data %>%
  filter(!grepl("U", Female_ID), !females_removed, mom_at_init,
         clutch_initiated_order == 1, visit_ends < last_day_fID) %>% 
  filter(!is.na(`visited in 4 days`)) %>%
  mutate(`visited in 4 days` = 
           ifelse(`visited in 4 days` == "Visited", 
                  "Visited\nwithin first 4 days", 
                  "Not visited\nwithin first 4 days")) %>%
  ggplot(aes(x=guard_duration, fill=ants)) +
  geom_histogram(binwidth = 4) +
  theme_publilia() +
  xlab("Guarding duration") +
  ylab("Number of females") +
  scale_y_continuous(breaks = c(0,4,8),
                     labels = label_number(accuracy = 1)) +
  scale_color_manual(values = c("black", "brown1")) +
  scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = 20, linetype="dashed", 
                color = "gray", size=1) +
  ggh4x::facet_grid2(vars(`visited in 4 days`, ants), 
                     strip = hist_strips) +
  theme(strip.background = element_rect(color = NA),
        strip.text = element_text(size = 12),
        legend.justification = c(1,1),
        legend.position = c(1,0.2),
        legend.text = element_text(size = 12)) +
  geom_text(data = hatch_text,
            aes(x = 20, y = Inf, label = label),
            hjust = -0.1, vjust = 0, color = "gray", fontface = "bold") +
  coord_cartesian(clip = "off"))
```

# 2023.07.22 - Using geom_ridgeline instead of freq_poly with alpha
Sarah asked to remove the stats text
```{r, fig.width=3, fig.height=2}
# Visualization: box plots with p-values
Fig1_v1D +
    geom_text(data=strats_cuts_v1.multcompLetters, 
              aes(x=vis_cat, y=-4, label=label, 
                  color=vis_cat%in%c("0","1 to 4")), 
              fontface="bold", size=6) +
  xlab("Day first visited (Days since initiation)") +
  ylab("Guard duration of initiating female") +
  theme(legend.position = "none")
```

>>>>

base_text <- expand.grid(ants = c("Ants removed", "Control"),
            `visited in 4 days` = c("Not visited", 
                                    "Visited")) %>%
  tibble()

hatch_text <- base_text %>%
  mutate(label = ifelse(ants=="Ants removed" & `visited in 4 days`=="Not visited", 
         "Hatch date", 
         ""))

(guarding_durations_plot_2023 <- strats_data %>%
  filter(!grepl("U", Female_ID), !females_removed, mom_at_init,
         clutch_initiated_order == 1, visit_ends < last_day_fID) %>% 
  filter(!is.na(`visited in 4 days`)) %>%
  mutate(`visited in 4 days` = 
           ifelse(`visited in 4 days` == "Visited", 
                  "Visited", 
                  "Not visited")) %>%
  ggplot(aes(x=guard_duration, fill=ants)) +
  geom_histogram(binwidth = 4) +
  theme_publilia() +
  xlab("Guarding duration") +
  ylab("Number of females") +
  scale_y_continuous(labels = label_number(accuracy = 1)) +
  scale_color_manual(values = c("black", "brown1")) +
  scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = 20, linetype="dashed", 
                color = "gray", size=1) +
  ggh4x::facet_grid2(vars(`visited in 4 days`, ants), 
                     strip = hist_strips) +
  theme(strip.background = element_rect(color = NA),
        strip.text = element_text(size = 12),
        legend.justification = c(1,1),
        legend.position = c(1,0.2),
        legend.text = element_text(size = 12)) +
  geom_text(data = hatch_text,
            aes(x = 20, y = Inf, label = label),
            hjust = -0.1, vjust = 0, color = "gray", fontface = "bold") +
  coord_cartesian(clip = "off"))
```

# 2023.09.31 Figure S4

## Original
```{r}

Fig1_v1A.gd <- Fig1_v1_data %>%
  filter(`visited in 4 days` == "Not visited") %>%
  ggplot(data = .) +
  geom_area(aes(x = guard_duration, y=..density.., color = ants, fill = ants), 
            alpha = 0.75, stat='bin', binwidth = 4, size = 0, position = "identity") + 
  theme_publilia() +
  xlab("Guarding duration") +
  ylab("Proportion\nof initiating females") +
  scale_color_manual(values = c("black", "brown1")) +
 scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = 20, linetype="dashed", 
                color = "gray", size=1)

Fig1_v1B.gd <- Fig1_v1_data %>%
  filter(`visited in 4 days` == "Visited") %>%
  ggplot(data = .) +
  geom_area(aes(x = guard_duration, y=..density.., color = ants, fill = ants), 
            alpha = 0.75, stat='bin', binwidth = 4, size = 0, position = "identity") + 
  theme_publilia() +
  xlab("Guarding duration") +
  ylab("Proportion\nof initiating females") +
  scale_color_manual(values = c("black", "brown1")) +
  scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = 20, linetype="dashed", 
                color = "gray", size=1)

Fig1_v1C.gd <- strats_data %>%
  filter(!females_removed, !mom_at_init,
         visit_ends >= initiation_date,
         visit_ends < last_day_fID) %>%
  ggplot(data = .) +
  geom_area(aes(x = guard_duration, y=..density.., color = ants, fill = ants), 
            alpha = 0.75, stat='bin', binwidth = 4, size = 0, position = "identity") + 
  theme_publilia() +
  xlab("Guarding duration") +
  ylab("Proportion\nof secondary visitors") +
  scale_color_manual(values = c("black", "brown1")) +
  scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = 20, linetype="dashed", 
                color = "gray", size=1)

## Histogram
```{r, fig.width=3, fig.height=4.5}
visit_colors <- rev(unique(ggplot_build(Fig1_v1D)$data[[1]]$fill))

hist_strips <- strip_nested(
     # Horizontal strips
     background_y = list(
       element_rect(fill = visit_colors[1]), 
       element_rect(fill = visit_colors[2]),
       element_rect(fill = "cornflowerblue"),
       element_blank(), element_blank(), 
       element_blank(), element_blank(),
       element_blank(), element_blank()),
     text_y = list(
       element_text(color = c("black")), 
       element_text(color = c("black")), 
       element_text(color = c("black")),
       element_blank(), element_blank(),
       element_blank(), element_blank(), 
       element_blank(), element_blank()),
     by_layer_y = FALSE
)

base_text <- expand.grid(ants = c("Ants removed", "Control"),
            cat4facet = c("Not visited", "Visited", "Visitor")) %>%
  tibble()

hatch_text <- base_text %>%
  mutate(label = ifelse(ants=="Ants removed" & cat4facet=="Not visited", 
         "Hatch date", 
         ""))

Figure_S4_data <- strats_data %>%
  mutate(is_not_visited = !grepl("U", Female_ID) &
           !females_removed &
           mom_at_init & 
           clutch_initiated_order == 1 &
           visit_ends < last_day_fID &
           `visited in 4 days` == "Not visited",
         is_visited = !grepl("U", Female_ID) &
           !females_removed &
           mom_at_init & 
           clutch_initiated_order == 1 &
           visit_ends < last_day_fID &
           `visited in 4 days` == "Visited",
         is_visitor = !females_removed &
           !mom_at_init &
           visit_ends >= initiation_date &
           visit_ends < last_day_fID,
         cat4facet = ifelse(is_visited, "Visited",
                            ifelse(is_not_visited, "Not visited",
                                   ifelse(is_visitor, "Visitor", NA))),
         cat4facet = factor(cat4facet, levels = c("Not visited", "Visited", "Visitor"))) %>%
  filter(!is.na(cat4facet))

(Figure_S4 <- Figure_S4_data %>%
  ggplot(aes(x=guard_duration, fill=ants)) +
  geom_histogram(binwidth = 4) +
  theme_publilia() +
  xlab("Guarding duration") +
  ylab("Number of females") +
  scale_y_continuous(labels = label_number(accuracy = 1)) +
  scale_color_manual(values = c("black", "brown1")) +
  scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = 20, linetype="dashed", 
                color = "gray", size=1) +
  ggh4x::facet_grid2(vars(cat4facet, ants), 
                     strip = hist_strips) +
  theme(strip.background = element_rect(color = NA),
        strip.text = element_text(size = 12),
        legend.justification = c(1,1),
        legend.position = c(1,0.2),
        legend.text = element_text(size = 12)) +
  geom_text(data = hatch_text,
            aes(x = 20, y = Inf, label = label),
            hjust = -0.1, vjust = 0, color = "gray", fontface = "bold") +
  coord_cartesian(clip = "off"))
```

# 2023.07.22 - Using geom_ridgeline instead of freq_poly with alpha
Sarah asked to remove the stats text
```{r, fig.width=3, fig.height=2}
# Visualization: box plots with p-values
Fig1_v1D +
    geom_text(data=strats_cuts_v1.multcompLetters, 
              aes(x=vis_cat, y=-4, label=label, 
                  color=vis_cat%in%c("0","1 to 4")), 
              fontface="bold", size=6) +
  xlab("Clutch age (days) when first visited") +
  ylab("Guard duration of initiating female") +
  theme(legend.position = "none")
```
